### 2019-01-14 papers/journal/intro.tex
Added new introduction to the paper

%\textbf{Components of the vesicle transport network}:
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane-bound organelles called compartments.
% 
These compartments exchange molecules amongst themselves in small membrane-bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target compartments for each vesicle is performed by an intricate network of molecules called the vesicle transport system~\cite{alberts2002molecular}. 
%
We call the set of internal compartments, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Vesicle traffic underpins nearly every aspect of eukaryotic cellular physiology, including
human cellular physiology.
%
Understanding how this system function is, therefore,
one of the central challenges of cell biology.
%
The high complexity of the problems~\cite{mani2016wine, mani2016stacking} in VTS restricts the use of traditional methods like simulation and statistical reasoning. 
%The high complexity of problem restricts the use of traditional methods like simulation and statistical reasoning and the requirement of precise analysis makes formal methods an ideal candidate. 
The requirement of precise analysis makes formal methods ideal candidate in this case.

First, we will give an overview of the VTS mechanism and then will sketch out a few interesting problems that are key to the understanding of the VTS.

\subsection{Vesicle traffic system}
\todo{SA:One para of BIO BACKGROUND related to VTS. Why they are Important history etc.}
%
%\textbf{Components of the vesicle transport network}:
The vesicle transport network is used by cells to transport molecules across compartments.
% 
The mechanism of vesicle transport can roughly be broken down into two big steps:
% budding of vesicles from the source compartment and their fusion with the target compartment.
%
the process of formation of a vesicle by budding from the membrane called as vesicle budding and the fusion of these transport vesicle with its target compartment called as vesicle fusion.

\todo{SA:Need a Diagram}Vesicle budding process involves series of steps, sorting and packaging of the molecules, membrane deformation and budding out from the compartment.
%
First the sorting of cargo molecules takes place by cytosolic adaptor molecules. 
%
Cargo adaptors bind to short linear stretches of amino acids in the cargo molecules called sorting signals and thereby sequester them for packaging. 
%
Then the different adaptor proteins act at different compartments, for eg., sec23/24 is required for packaging vesicles that form at the ER, the AP2 adaptor is used for packaging vesicles at the plasma membrane, etc.
%
Thereafter, adaptors are recruited to the source compartments through interactions with the lipids that form the source compartment's membrane, and interactions with Arf/Sar GTPases which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}.

The next step is membrane deformation, which involves cytosolic coat proteins.
Coat proteins are recruited to the forming vesicle through their interaction with adaptor proteins. 
%
These are naturally curved proteins, and their assembly on membranes leads to membrane deformation.
%
A single kind of coat protein can bind to multiple kinds of adaptors, thereby allowing the packaging of multiple types of cargo within the same vesicle. 
%
Clathrin, COP2 and COP1 are the three major kinds of coat proteins in eukaryotic cells. 
%
The clathrin coat operates at multiple compartments, and can interact with multiple adaptors proteins, whereas the COP2 and COP1 coat are only involved in vesicle formation at a single compartment; the ER and the Golgi respectively~\cite{faini2013vesicle}.

The final step is budding of the vesicles from the source compartment.
% by dynamin.
%	The final step in vesicle formation is its pinching off from the source compartment.
	%
This process requires ATP and is performed by the cytosolic dynamin proteins, which are recruited to the vesicle neck by the coat proteins~\cite{cocucci2014dynamin}.
%
The transport vesicle is now ready for the transportation.

Vesicle fusion involves two types of events, recognition of the correct target compartment and fusion of the transport vesicle with it's target compartment.

	 \textbf{Recognition of the correct target compartment} is a complex process involving Rab
	GTPases and tethering complexes.
	% 
	 Rab GTPases are small proteins which occur in two forms: the membrane associated and active
		GTP bound form, and the cytosolic and inactive GDP bound forms. 
		%
		Different compartments of the
		cell are associated with a different type of Rab protein. In their active form, Rabs recruit many
		downstream proteins, including tethering complexes, fusion regulators, motor proteins, sorting
		adaptors, etc, to the membrane they are associated with. 
		%
		The proteins that are recruited by active
		Rabs are called Rab effectors.
		%
		Rab activity is controlled by two proteins: GEFs turn Rabs on and
		GAPs turn them off. 
		%
		Each Rab has its own specific GEF and GAP. In some cases, GEFs and GAPs
		themselves can be Rab effectors, thus generating feed-forward and feedback loops. 
		%
		In their cytosolic, inactive form, Rabs are complexed with GDI proteins, and it is in this form that Rabs are presented to membrane bound GEFs for activation~\cite{muller2018molecular}

	 \textbf{Tethering complexes} sequester vesicles to their target compartments and regulate SNARE
		proteins, which enable vesicle fusion. 
		%
		Tethers are believed to bind to SNAREs and function as
		chaperones for SNARE complex assembly which is the ultimate step in membrane fusion. 
		%
		There are two types of tethering complexes: multi subunit tethers and long coiled-coil tethers:
%		
 Coiled coil tethers are long, single subunit proteins. 
			%
			These tethers are believed to bind the
			vesicle on one end and the target compartment on the other end, thereby bridging the two before
			fusion.
			%
			A clear mechanism for action of these tethers is missing, but based on their interactions with
			other proteins of the traffic machinery, some hypotheses as to their mechanism of action have been
			put forth. 
			%
			Golgins, coiled coil tethers that are anchored at the Golgi by a transmembrane domain, contain a large number of binding sites for Rab GTPases. 
			%
			Thus, golgins might capture vesicles by
			binding vesicle-associated Rab proteins. 
			%
			Golgins also contain domains which can sense membrane
			curvature, which they could be using to recognize vesicles. 
			%
			They might also be interacting directly
			with vesicle SNAREs~\cite{baker2016chaperoning}.
		%	
			Multi-subunit tethering complexes (MTCs) are composed of three or more different
			subunits.
			%
			MTCs are known to interact with vesicle coat proteins, Rab GTPases, SNAREs and SM proteins.
			%
			For example, the HOPS complex, which is required for homotypic (fusion of two identical
			membranes) and heterotypic fusion in the endo-lysosomal system, was shown to tether membranes
			through its interactions with the membrane-associated Rab GTPase Ypt7, acidic phospholipids and
			SNAREs. 
			%
			The HOPS complex binds both individual SNAREs and SNARE complexes. It also
			seems to protect assembling trans-SNARE complexes from premature disassembly. 
			%
			Another example is given by the Dsl1 complex, which is anchored to the ER membrane through interactions
			with t-SNAREs. 
			%
			At the other end, the Dsl1 complex contains multiple binding sites for COPI
			(vesicle coat that is present on vesicles produced by the Golgi). 
			%
			This structure suggests that the Dsl1
			complex functions as a tether connecting COPI-coated vesicles to their target organelle, the ER~\cite{baker2016chaperoning}.

	 \textbf{Fusion of vesicle with the target compartment} is brought about by SNARE proteins. 
	%
	SNAREs
	are defined by a 60- to 70-residue SNARE motif. Most SNAREs are anchored to the membrane by
	their C-terminal transmembrane helices. 
	%
	The formation of a productive SNARE complex requires
	the formation of a four-helix bundle containing four different SNARE motifs. Usually, one of these
	motifs is contributed by the SNARE on the vesicle membrane (v-SNAREs), and the other three
	motifs by SNAREs on the membrane of the tartget compartment (t-SNAREs). 
	%
	Alternatively, they
	can be defined by their amino acid sequences as R-, Qa-, Qb-, and Qc-SNAREs. A few SNARE
	proteins, such as SNAP25, contain both Qb- and Qc-SNARE motifs.
	%
	Most v-SNAREs are RSNAREs, and most t-SNAREs are Q-SNAREs~\cite{yoon2018snare}.

Many SNAREs also have N-terminal domains that regulate SNARE complex assembly and/or
interact with other parts of the vesicle fusion machinery. 
%
Qa-SNAREs have an amino terminal
domain called the Habc domain, which can fold back on the SNARE domain and hold the QaSNARE
in an inactive state. 
%
This inactivated state is stabilized by interactions with SM proteins~\cite{yoon2018snare}.

A subset of R-SNAREs (the longin SNAREs) possess an amino-terminal domain called the longin domain. 
%
In addition to regulation of SNARE activity as in the case of the Qa-SNARE Habcdomains,
these longin domains also regulate the localization of longin SNAREs through their
interaction with adaptor proteins~\cite{daste2015structure}.

Different vesicle-target compartment pairs in the cell are associated with unique SNARE
complexes.
%
Membrane fusion converts the trans-SNARE complex (v- and t- SNAREs on opposite
membranes) into a cis-SNARE complex (both v- and t- SNAREs on the same membrane).
Disassembly of cis-SNARE complexes requires Sec17 and Sec18 (the yeast SNAP and NSF
protein, respectively). 
%
This process requires energy which is released by the hydrolysis of ATP.

\textbf{SNAREs are regulated by SM proteins}, which have three modes of action: SM proteins can hold
the Qa-SNAREs in an autoinhibited state and prevent or postpone their assembly into SNARE
complexes.
%
Secondly, some SM proteins have been seen to act as a template upon which a halfzippered
complex between the Qa- and R-SNAREs (an early SNARE complex intermediate) can
form. 
%
In this mode, by choosing an R-SNARE located on the opposite membrane, SM proteins
could be imposing a filter inhibiting the formation of futile cis-SNARE complexes and promoting
trans-SNARE complexes.
%
Finally, SM proteins bind the four-helix bundles formed by assembled
SNARE complexes.
%
This mode of binding of SM proteins might protect assembled SNARE
complexes from premature disassembly by NSF and SNAPs and/or might stimulate fusion directly~\cite{baker2016chaperoning}.
%
SM proteins have also been shown to interact with tethers~\cite{yoon2018snare}.

\textbf{Major paths in the vesicle transport network}:
Molecules traverse the cell in a series of vesicles, and there are two major such routes of transport in all eukaryotic cells.
%
	The secretory route takes proteins from the ER, their site of production, to the plasma membrane, from where they are secreted out of the cell.
	%
	 Cargo proteins leave the ER in COP2 coated vesicles which fuse with the Golgi apparatus.
	 %
	  The golgi apparatus is where proteins get modified, for example by the addition of carbohydrate side-chains. 
	  %
	  Subsequently, proteins destined for secretion are packaged into clathrin coated vesicles which fuse with the plasma membrane, thus releasing their contents to the outside of the cell\cite{alberts2002molecular}.
	  %
 The endocytic routetakes proteins from the outside of the cell through the plasma membrane to the endocytic compartments, where they are digested. 
 %
 Cargo from the outside of the cell is taken in in the form of clathrin coated vesicles which form at the plasma membrane. 
 %
 These vesicles fuse amongst themselves in a process called homotypic fusion to generate the early endosomal compartments. 
 %
 Early endosomes undergo changes in their composition due to ongoing vesicle traffic and maturee into late endosomes. 
 %
 This maturation is concomitant with a switch the Rab protein they are associated with; the Rab5 associated early endosomes mature into Rab7 associated late endosomes\cite{rink2005rab}. Late endosomes then fuse with lysosomes, where all its contents get digested\cite{pryor2009delivery}.
%
Other paths are used for cross talk between the secretory and the endocytic routes. For example, vesicles are sent from the TGN (Trans-Golgi network) to the early and the late endosomes to transport enzymes, and vesicles are sent back from the endocytic compartments to the TGN to recycle sorting receptors\cite{progida2016bidirectional}. Also recently, evidence for unconventional vesicle-mediated secretory routes have been found, which bypass the Golgi. The molecules involved in these routes are as yet unknown\cite{nickel2018unconventional}. 
%

@@

In this section, we will abstract from the biological description of the VTS and represent the whole network as an annotated transport graph. 
%
The constraints imposed by cell biology is incorporated in the  annotations of the graph.

\textbf{The cell as a transport graph:} 
%We consider a cell to be a collection of compartments (nodes) and vesicles (edges), thus defining a transport graph. Every compartment or vesicle has a set of molecular labels, such as SNARE proteins, associated with it.
The cell can be abstractly represented as an annotated graph. 
Every compartment in the cell can be represented as a node in the graph. 
%
The set of molecules present in the compartment, for example, SNARE proteins associated with it, are represented as a label to the corresponding nodes.
%
The label of each node will be unique molecules present on the compartment, i.e we abstract over the quantity of the molecule present in a compartment.
%This represents the flux of the molecule type present at that compartment.
%
The target vesicle flowing from source compartment to the target compartment is represented by a labeled directed edge. 
% 
The label represents the associated flux of all molecular types carried by the corresponding vesicle.
%
%

\textbf{Steady state:} 
The vesicle transfer will change the molecular composition (distinct molecule count) on both the source compartment and the target compartment. 
%
In our abstracted model, we will assume that cell is in a steady state. 
%
The compartment's composition does not evolve over the short times' scales.
%
Therefore, the incoming and outgoing flux is balanced for each of the molecular types at each compartment.
%
This abstraction is supported by the cell biology as on the scale of seconds and minutes the molecular composition actually remains the same. 
%
In this paper, we refer to the steady state of the cell as stability condition over the annotated graph.

@@

\textbf{Vesicle fusion:}
%Based on the earlier discussion of fusion, the vesicle targeting is driven by molecular interactions.  
%
%Particularly, molecules composition (SNARE proteins etc) present on the budded vesicle determines it's properties, which 
% 
%The molecular composition and hence the properties of the transfer vesicle 
%is the crucial factor to which target compartment the vesicle will fuse to.
%
%For any given pair of a vesicle and a compartment, the SNARE proteins present on both influence the fusion of the vesicle to that compartment.
%  
%Biophysically, fusion of a vesicle to the target compartment requires a direct physical interaction between at least one SNARE type on the vesicle and one SNARE type on the compartment.
%
%Once a vesicle has budded out of the source, the molecules it carries determine its properties. In particular, for any given pair of a vesicle and a compartment, the set of SNARE proteins that label the former and latter influence whether the vesicle will fuse to that compartment. Biophysically, fusion requires a direct physical interaction between at least one SNARE type on the vesicle and one SNARE type on the compartment. 
%
%SNAREs are of two types (known as Q and R in the cell biology iterature) and 
Vesicle fusion requires a pairing of a Q-SNARE with an R-SNARE between transfer vesicle and target compartment. 
%For any given pair of a vesicle and a compartment, vesicle fusion requires a pairing of a Q-SNARE with an R-SNARE on 
%the SNARE proteins present on both influence the fusion of the vesicle to that compartment.
%
%The list of molecular pairs that can drive a fusion event is given in a pairing matrix between Q and R SNAREs. 
%
Only a particular pair of Q and R SNAREs are allowed to fuse with each other.
%
Without loss of generality, we assume equal numbers of Q and R SNARE types.
%
Therefore, we can abstract the underlying cell biology by labeling the nodes and edges with an equal number of Q and R SNAREs. 
% equal number of Q and R SNAREs as a part of node and edge label.
%
%
Given a relation of all allowed fusion SNARE pairings, we can computationally determine whether a particular transfer vesicle will fuse to a particular target compartment (the edge between two compartments) based on the above condition.

\textbf{Molecular regulation:} 
Fusion takes place only if the SNARE types involved in the vesicle and compartment must both be in an active state.
%We assume that for fusion to occur, the pair of 
%
The activity of these SNAREs is dependent on the presence of other molecules on the vesicle or compartment, respectively.
%
In our abstract model, we create a set of variations of different kind of molecular regulations.
%
Most generally, the activity state of a given SNARE can be a Boolean function of all the molecular types on a compartment or vesicle. 
%
We have also tested \cite{shukla} a particularly simple regulation mechanism in which two SNAREs that can pair to drive fusion to inhibit one another; this is the \textit{pairing inhibition}. 
%
This is motivated by the idea that pairing must generate an inactive bi-molecular complex.
%

@@

\subsection{Interesting properties of VTS}
%\textbf{Properties of a VTS that satisfies all cell-biological constraints:} 
Suppose we are given a particular transport graph, a particular labeling of all the compartments and edges, a particular fusion pairing matrix, and a particular regulatory model. This information is then sufficient to check the following properties, which summarise the cell-biological constraints described above:
\begin{enumerate}
	\item We can determine which molecules are active on every compartment or vesicle.
	\item For every vesicle fusing to a compartment, we can determine whether there exists an active pair (one molecule on the vesicle, one on the compartment) which drives that fusion event.
	\item For every vesicle-compartment pair where the vesicle does not fuse to the compartment, we can verify that there is no pairing of active molecules on the
	vesicle and compartment that could drive their fusion.
	\item We can verify that every molecular type entering a compartment also leaves the compartment, and also that every molecular type entering a set of compartments also leaves that set. This is the steady state condition. In the biological literature this is often referred to as “homeostasis” and is a widespread and well-accepted assumption about cellular behaviour, at least over timescales of minutes to hours \cite{mani2016stacking}.
\end{enumerate}

The biological problem often boils down to such an analysis. A cell biologist might determine which molecules flow between which set of compartments, and biochemical experiments could be used to see how these molecules activate one another. We can then ask: is this a complete and consistent description? That is, do all the required properties listed above hold, given what the experimentalists have told us? It is often the case that biological data is missing. For example, only a few of the dozens of molecules involved in real VTSs have been mapped out. Therefore, it is extremely likely that the description provided by the cell biologist is incomplete. We can use our model to find out which properties have failed to hold, and thus prescribe new experiments in order to fill in the missing information.

Can we find a simple test to see whether any information is missing, given the experimental data? We have shown that graph k-connectedness furnishes precisely such a test \cite{shukla}. If the data provided by cell-biologists, suitably represented as a graph, does not have a certain degree of connectivity, this implies that some biological data has been missed. (The converse is not true: even if the required connectivity does hold, there might of course be more information missing.)

Our result about k-connectedness being a useful test of missing information \cite{shukla} was obtained using SAT solvers for graphs up to a certain size, and a certain number of molecules. This was due to limitations in how we encoded the problem. Here we present a much more natural encoding that allows our result to be extended to graph sizes and molecule numbers that are typical of those found in real cells.

%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}
%
Since a VTS is a transport graph, it is but natural to formally model
VTS as graphs (as used in computer science) with their nodes denoting
compartments and labeled edges denoting transport vesicles with labels
denoting the set of molecules being transported. The pairing mechanism
can be represented as matching tables over sets of molecules.
% i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met. 
%
Given such a graph model of VTSs and their properties, such as
stability condition and fusion rules, can be formally defined as
constraints over graphs and uninterpreted Boolean functions.
%
% Note that the formulas define among other things constraints over
% paths between nodes in the graph model of VTS. Similarly, one can
% define fusion rules as constraints over the boolean functions modeling
% the regulations.  
\
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.

@@

%A VTS is {\em $k$-connected} if every pair of compartments remain reachable after dropping $k-1$ vesicles.
%
%This property of VTSs has been considered informative and studied by~\cite{shukla}.
%
%Here we have build an {\em efficient} tool that studies properties of VTSs that are not $k$-connected, from some $k$.

\subsection{Synthesis of VTS:}
Although vesicle-mediated traffic was discovered decades ago~\cite{wells2005discovery}, the picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident organelles within the cell, and new vesicle routes are being discovered every year ~\cite{nickel2018unconventional,weill2018toolbox}. 
%
Completing the vesicle traffic network is a difficult task for the following reasons: 
\begin{enumerate}
	\item The core vesicle transport network, which consists of a secretory and an endocytic route, is conserved across all eukaryotes, but the traffic network in different organisms \cite{richardson2015evolutionary,nishimoto2009differential,barlow2017seeing}, and even in different cell types within an organism can be different ~\cite{stoops2014trafficking,zhou2015arp2}.

	\item Although the basic traffic machinery is the same for all vesicle fusion events, the details of regulation can be different~\cite{davletov2007regulation,di2010calcium}.

	\item Behaviour of traffic molecules in vitro is different from their behaviour inside cells ~\cite{furukawa2014multiple}.

	\item Molecules can have redundant routes within the cell ~\cite{shimizu2014compensatory,nakatsukasa2014nutrient}.

>>>>

A characteristic feature of eukaryotic cells is the presence of multiple internal membrane bound compartments. These compartments exchange molecules amongst themselves in small membrane bound packets called vesicles ~\cite{alberts2002molecular}.  The molecular interaction network that controls the vesicle transport system involves tens of kinds of molecules which regulate each others activities and modes of action. Due to this, obtaining a complete understanding of the vesicle transport system through physical experimentation alone is a hard problem.
%
The high complexity of problems in VTS also restricts the use of traditional methods like simulation and statistical reasoning~\cite{mani2016wine, mani2016stacking}. 
%The high complexity of problem restricts the use of traditional methods like simulation and statistical reasoning and the requirement of precise analysis makes formal methods an ideal candidate. 
In this paper, using formal methods, we solve the problem of completing incomplete VTS graphs while also considering constraints due to molecular interactions, thereby demonstrating the aptitude of such methods.

First, we give an overview of the molecular mechanism of the VTS and then we sketch out a few interesting problems that are key to an understanding of the VTS.

\subsection{Vesicle traffic system}
Vesicles were discovered decades ago, in the early 1950s, and major molecules of the vesicle traffic system and their functions were discovered soon after~\cite{wells2005discovery}. These molecules can roughly be categorized as budding molecules, whose role is in the formation and budding of transport vesicles from the source compartment, and fusion molecules, which enable fusion of transport vesicles with the target compartments. Below, we give a brief description of the sequence of steps involved in the vesicle transport process. Through this description, we wish to provide a glimpse of the kind of molecular regulatory interactions that we encode in our model. 
\begin{figure}
	\includegraphics[width=\textwidth]{ankitspaper_schem1.png}
	\label{schematic}
	\caption{Schematic of the vesicle transport system. Steps involved in vesicle transport between two compartments; a source and a target compartment are shown. The symbols used for different molecules involved in vesicle traffic are explained in the key. At the source compartment, transmembrane cargo proteins are sorted by cytosolic adaptor proteins. Here, the different colours indicate different types of cargo and adaptor proteins. Coat proteins are recruited to the forming vesicle by adaptor proteins. Note here that a single type of coat protein can interact with multiple types of adaptor proteins. Subsequently, the vesicle is pinched off at its neck by dynamin. Once the vesicle has left the source compartment, the coat and adaptor proteins fall off and the vesicle is recognized and bound by a tethering complex whose one end is attached to the target compartment. At the target compartment, the tethered vesicle undergoes fusion, thereby releasing its contents into the target compartment. }
\end{figure}

\textbf{Vesicle budding} involves sorting and packaging of cargo molecules, membrane deformation and budding of the vesicle from the source compartment~\ref{schematic}.
%
Cargo sorting is performed by cytosolic adaptor proteins. 
Adaptors are recruited to the source compartment through interactions with the lipids that form the compartment's membrane, and interactions with molecular switches called Arf/Sar which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}. Then these adaptors bind to cargo molecules and thereby sequester them for packaging. 
%
Different adaptor proteins act at different compartments, for eg., vesicle formation at the ER requires the adaptor sec23/24, and AP2 adaptor is used for packaging vesicles at the plasma membrane.
%

The next step is membrane deformation, which involves cytosolic coat proteins. Coat proteins are recruited to the forming vesicle through their interaction with adaptor proteins. These are naturally curved proteins, and their assembly on membranes leads to membrane deformation.
A single kind of coat protein can bind to multiple kinds of adaptors, thereby allowing the packaging of multiple types of cargo within the same vesicle. 
%
There are three major types of coat proteins; Clathrin, COP2 and COP1. These act on different compartments; the clathrin coat operates at multiple compartments, and can interact with multiple adaptors proteins, whereas the COP2 and COP1 coat are only involved in vesicle formation at a single compartment; the ER and the Golgi respectively~\cite{faini2013vesicle}.

Finally, transport vesicles bud out from the source compartment. This process requires ATP and is performed by the cytosolic dynamin proteins, which are recruited to the vesicle neck by the coat proteins~\cite{cocucci2014dynamin}.
%
The vesicle is now ready for transport and subsequent fusion with a target compartment.

\textbf{Vesicle fusion involves two steps:} recognition of the correct target compartment and fusion of the transport vesicle~\cite{schematic}.

Target compartment recognition involves molecular switches called Rabs and  tethering complexes. Distinct compartments of the cell are associated with distinct Rabs. For example, Rab5 is associated early endosomes and  Rab7 with late endosomes\cite{rink2005rab}.  Rabs are master regulators; in their active form Rabs recruit many downstream proteins, including tethering complexes, fusion regulators, motor proteins, etc, to their compartment. ~\cite{muller2018molecular}

Tethering complexes sequester vesicles to their target compartments by binding to the vesicle on one end and the target compartment on the other end, thereby bridging the two before fusion. They are also believed to function as chaperones for assembly of SNARE protein complexes, which enable vesicle fusion. 
A clear mechanism for action of these tethers is missing, but evidence from experiments indicate that tethers bind to vesicles through molecular interactions with vesicle coat proteins, vesicle associated Rabs, SNAREs and SNARE regulators. They can also sense membrane curvature (vesicles are highly curved) ~\cite{baker2016chaperoning}.

The final step in vesicle transport is fusion of vesicle with the target compartment.This is brought about by SNARE proteins. 
There are four kinds of SNAREs: R, Qa, Qb, and Qc; and vesicle fusion requires formation of a SNARE complex containing one of each type. In all known cases, one SNARE protein of the complex is contributed by the vesicle and the other three SNAREs by the target compartment (A few SNARE proteins, such as SNAP25, contain both Qb- and Qc-SNARE motifs)~\cite{yoon2018snare}.

\todo{schematic of snare regulatory network}
In keeping with their central role in vesicle transport,both the activity and the localization of SNARE proteins are regulated. Notable regulators are Sec17/18 and the SM proteins.
%
\newline Sec17/18 bring about disassembly of SNARE complexes post-fusion in order that these SNAREs may be reused. 
\newline SM proteins have multiple modes of SNARE regulation: 
\begin{itemize}
	\item SM proteins can hold
	the Qa-SNAREs in an inhibited state and prevent or postpone their assembly into SNARE
	complexes.
	%
	\item Some SM proteins can act as a template upon which an early SNARE complex intermediate can
	form.
	%
	\item Finally, SM proteins bind the fully assembled
	SNARE complexes to protect them from premature disassembly by Sec17/18~\cite{baker2016chaperoning}.
	%
	\item SM proteins have also been shown to interact with tethers~\cite{yoon2018snare}.
\end{itemize}

\textbf{Major paths in the vesicle transport network}:
\todo{schematic of intracellular compartments and  major routes}
Molecules traverse the cell in a series of vesicles, and there are two major routes of vesicle transport in all eukaryotic cells.
%
\begin{itemize}
	\item The secretory route takes proteins from the ER (their site of production) to the plasma membrane( from where they are secreted out of the cell) via the Golgi apparatus\cite{alberts2002molecular}.
	%
	\item The endocytic route takes proteins from the outside of the cell, through the plasma membrane, to the endocytic compartments and the lysosome where they are digested. 
\end{itemize}
Other paths are used for cross talk between the secretory and the endocytic routes. For example, vesicles are sent back and forth between the Trans-Golgi network (secretory route) to the early and the late endosomes(endocytic route)\cite{progida2016bidirectional}. Also, recently, unconventional vesicle-mediated secretory routes have been found, which bypass the Golgi apparatus. The molecules involved in these routes are as yet unknown\cite{nickel2018unconventional}.

%

@@

In this section, we will abstract from the biological description of the VTS and represent the whole network as an annotated transport graph. 
%
The constraints imposed by cell biology are incorporated in the  annotations of the graph.

\textbf{The cell as a transport graph:} 
%We consider a cell to be a collection of compartments (nodes) and vesicles (edges), thus defining a transport graph. Every compartment or vesicle has a set of molecular labels, such as SNARE proteins, associated with it.
The cell can be abstractly represented as an annotated graph. 
Every compartment in the cell can be represented as a node in the graph. 
%
The set of molecules present in the compartment, for example, SNARE proteins associated with it, are represented as a label to the corresponding nodes.
%
The label of each node will be unique molecules present on the compartment, i.e we abstract over the quantity of the molecule present in a compartment.
%This represents the flux of the molecule type present at that compartment.
%
The transport vesicle flowing from source compartment to the target compartment is represented by a labeled directed edge. 
% 
The label represents the associated flux of all molecular types carried by the corresponding vesicle.
%
%

\textbf{Steady state:} 
The vesicle transfer will change the molecular composition (distinct molecule count) on both the source compartment and the target compartment. 
%
In our model, we assume that the cell is in steady state. As a further simplification, we also assume that the incoming and outgoing flux is balanced for each of the molecular types at each compartment. 
Indeed, it is well accepted that on the scale of minutes and hours the molecular composition remains the same. 
%
In this paper, we refer to the steady state of the cell (refered as \textit{homeostasis} in biological literature) as the stability condition over the annotated graph.

@@

\textbf{Vesicle fusion:}
%Based on the earlier discussion of fusion, the vesicle targeting is driven by molecular interactions.  
%
%Particularly, molecules composition (SNARE proteins etc) present on the budded vesicle determines it's properties, which 
% 
%The molecular composition and hence the properties of the transfer vesicle 
%is the crucial factor to which target compartment the vesicle will fuse to.
%
%For any given pair of a vesicle and a compartment, the SNARE proteins present on both influence the fusion of the vesicle to that compartment.
%  
%Biophysically, fusion of a vesicle to the target compartment requires a direct physical interaction between at least one SNARE type on the vesicle and one SNARE type on the compartment.
%
%Once a vesicle has budded out of the source, the molecules it carries determine its properties. In particular, for any given pair of a vesicle and a compartment, the set of SNARE proteins that label the former and latter influence whether the vesicle will fuse to that compartment. Biophysically, fusion requires a direct physical interaction between at least one SNARE type on the vesicle and one SNARE type on the compartment. 
%
%SNAREs are of two types (known as Q and R in the cell biology iterature) and 
Vesicle fusion requires a pairing of 3 Qsnares and 1 Rsnare that are distributed between the transfer vesicle and the target compartment. 
%For any given pair of a vesicle and a compartment, vesicle fusion requires a pairing of a Q-SNARE with an R-SNARE on 
%the SNARE proteins present on both influence the fusion of the vesicle to that compartment.
%
%The list of molecular pairs that can drive a fusion event is given in a pairing matrix between Q and R SNAREs. 
%
Not all sets of Q and R SNAREs are allowed to fuse with each other.
%
In the model, we assume that the cell has an equal number of Q and R SNARE types.
%
Therefore, we can abstract the underlying cell biology by labeling the nodes and edges with an equal number of Q and R SNAREs. 
% equal number of Q and R SNAREs as a part of node and edge label.
%
%
Given a relation of all allowed fusion SNARE pairings, we can computationally determine whether a particular transfer vesicle will fuse to a particular target compartment (the edge between two compartments) based on the above condition.

\textbf{Molecular regulation:} 
Fusion takes place only if the SNARE types involved in the vesicle and compartment are both in an active state.
%We assume that for fusion to occur, the pair of 
%
The activity of these SNAREs is dependent on the presence of other molecules on the vesicle or compartment, respectively.
%
In our abstract model, 
we create a hierarchy of models for varying degrees of regulation of SNARE activity.
%we create a set of variations of different kind of molecular regulations.
%
Most generally, the activity state of a given SNARE can be a Boolean function of all the molecular types on a compartment or vesicle. 
%
We have also tested \cite{shukla} a particularly simple regulation mechanism in which two SNAREs that can pair to drive fusion to inhibit one another; this is the \textit{pairing inhibition}. 
%
This is motivated by the idea that pairing must generate an inactive bi-molecular complex.
%

@@

\subsection{Graph connectivity is an interesting and accessible property of the VTS}
%\textbf{Properties of a VTS that satisfies all cell-biological constraints:} 
Interactions between the molecules of the VTS together produce a functional vesicle transport network, in which molecules flow between intracellular compartments using specified paths. The molecules of the VTS, their interactions and modes of regulation have been worked out to a large extent through cell-biological experiments. But we do not know if this information is complete. In order to assess how completely we know the VTS, we need easily readable benchmarks.

\todo{SA: Need more text on the importance and relevance of PROPRTY OF CONNECTIVITY}
One such benchmark is provided by our result about connectivity of the VTS graph. Graph connectivity is defined as the minimum number of edges that need to be removed from an undirected graph to render it disconnected. For example, in a directed cyclic graph, every node has at least one incoming and one outgoing edge, at least 2 edges need to be removed in order to disconnect the graph. Therefore the underlying undirected graph is 2-connected. 
From our result, for a given set of biomolecular interactions, for a cell that is in steady state, we expect the VTS graph to have a certain k-connectivity. When regulation is given by arbitrary Boolean rules,   we find that VTS graphs must be 3-connected. And when we add the constraint of the specific form of regulatory interactions seen between VTS molecules, we expect vesicle transport graphs to be 2-strongly connected [28].

\todo{SA: We need not specify Synthesis here. Only connectivity use next subsection to build this. SM: but this is the para where we explain why connectivity is interesting an biological property}
The vesicle transport graphs available today for eukaryotic organisms are incomplete themselves. 
%
In this situation, we use our result as a tool to predict missing edges which satisfy the graph connectivity constraint. 
%
These predictions are intended to be used as guides for future experiments. 
%
A mismatch between our predictions and experiments would indicate that our current understanding of the VTS molecules is necessarily incomplete. 
%
Note that, on the other hand, if our predictions are experimentally confirmed, it does not imply that our understanding of the VTS is complete.

%\subsection{Modeling and Symbolic Analysis of VTS: An Overview}
\label{subsec:graphmodel}
%

% i.e., formally as a boolean function that given requisite labels of nodes and vesicles that returns true if the required regulations are met. 
%

%
% Note that the formulas define among other things constraints over
% paths between nodes in the graph model of VTS. Similarly, one can
% define fusion rules as constraints over the boolean functions modeling
% the regulations.  
\
% For example, the steady state condition, described informally earlier, can be defined by the formula shown in Listing~1.1.

@@

%A VTS is {\em $k$-connected} if every pair of compartments remain reachable after dropping $k-1$ vesicles.
%
%This property of VTSs has been considered informative and studied by~\cite{shukla}.
%
%Here we have build an {\em efficient} tool that studies properties of VTSs that are not $k$-connected, from some $k$. 
\subsection{Synthesis of VTS }
The current picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident compartments within the cell, and new vesicle routes are being discovered every year ~\cite{nickel2018unconventional,weill2018toolbox}. 
%
Completing the vesicle traffic network is a difficult task for the many important reasons: 
\begin{enumerate}
	\item The core vesicle transport network is conserved across eukaryotes, the traffic network in different organisms \cite{richardson2015evolutionary,nishimoto2009differential,barlow2017seeing}, and even in different cell types within an organism can be different ~\cite{stoops2014trafficking,zhou2015arp2}.

	\item Although the basic molecular machinery is the same for all vesicle fusion events, the details of regulation can be different~\cite{davletov2007regulation,di2010calcium}.

	\item Molecules can have redundant routes within the cell ~\cite{shimizu2014compensatory,nakatsukasa2014nutrient}.
	\item Behaviour of traffic molecules in vitro is different from their behaviour inside cells ~\cite{furukawa2014multiple}.
